---
title: "Network analysis in r"
author: "Mburu"
date: "6/19/2020"
output: github_document: null
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Creating an igraph object

Here you will learn how to create an igraph 'object' from data stored in an edgelist. The data are friendships in a group of students. You will also learn how to make a basic visualization of the network.

Each row of the friends dataframe represents an edge in the network.
```{r}
library(tidyverse)
library(data.table)
friends <- fread("friends.csv")

```

```{r, fig.width= 8}
# Load igraph
library(igraph)

# Inspect the first few rows of the dataframe 'friends'
head(friends)

# Convert friends dataframe to a matrix
friends.mat <- as.matrix(friends)

# Convert friends matrix to an igraph object
g <- graph.edgelist(friends.mat, directed = FALSE)


# Make a very basic plot of the network
plot(g)


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
