---
title: "Introduction to the Bioconductor"
author: "Mburu"
date: "5/27/2020"
output:
  github_document:
    df_print: paged
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Current version is using bioc manager
# Bioconductor version

In the video, you learned about the Bioconductor project. 
One advantage of this fantastic resource is its continuous improvements,
reflected in scheduled releases. Hence, checking the current version is 
important for the reproducibility of your analysis.

Important note: For this course, you will be using the BiocInstaller 
package to install and check package versions, as we will be using 
Bioconductor version 3.6. For details on the latest Bioconductor installation 
instructions see Bioconductor install page.

The package BiocInstaller has been installed using 
source("https://bioconductor.org/biocLite.R"). Remember, you can 
check the Bioconductor version using the function biocVersion() 
from the BiocInstaller package.


```{r}
# Load the BiocInstaller package
#install.packages("BiocManager")

library(BiocManager)

# Check R version
#version
sessionInfo()

# Explicit syntax to check the Bioconductor version
BiocManager::version()
```


## BiocLite to install packages
As example to install packages we will use the BSgenome package.
This is a data package which contains representations of several genomes.
When you install a new package it usually calls installations for its 
dependencies that takes some time, hence we have already installed BSgenome for you to use on this exercise.

Note: Using Bioconductor version 3.7 or earlier, packages can be installed using the following code:

# Using Bioconductor version 3.6
source("https://bioconductor.org/biocLite.R")
biocLite("BSgenome")

# Load the package
library(BSgenome)
For versions 3.8 or higher, you will use the BiocManager instead.

For this exercise, you should not use source() or biocLite() to install BSgenome.
This has already been done for you.

You can load the package using library(packageName) and check 
the package's version using the function packageVersion("packageName")

```{r}
#install("BSgenome")
# Load the BSgenome package
library(BSgenome)

# Check the version of the BSgenome package
packageVersion("BSgenome")



```

# S4 class definition

We will use the class BSgenome, which is already loaded for you.

Let's check the formal definition of this class by using the 
function showClass("className"). Check the BSgenome class results 
and find its parent classes (Extends) and the classes that inherit from it (Subclasses).

* Well done! The BSgenome is a powerful class and inherits from 
GenomeDescription, which you will see later on, and has MaskedBSgenome as subclass.

```{r}
showClass("BSgenome")
```


# Interaction with classes
Let's say we have an object called a_genome from class BSgenome.
With a_genome, you can ask questions like these:

# What is a_genome's main class?
class(a_genome)  # "BSgenome"

# What is a_genome's other classes?
is(a_genome)  # "BSgenome", "GenomeDescription"

# Is a_genome an S4 representation?
isS4(a_genome)  # TRUE
If you want to find out more about the a_genome object,
you can use the accessor show(a_genome) or use other specific 
accessors from the list of .S4methods(class = "BSgenome").

* Keep up the good work! You can now check other objects and 
investigate if they are S4 objects, their classes, and their
accessors. Remember, you can use .S4methods() or showMethods() 
to check the accessors list of a class or a function.


```{r}
# Investigate the a_genome using show()
# Read a_genome
library(BSgenome.Scerevisiae.UCSC.sacCer3)
a_genome <- BSgenome.Scerevisiae.UCSC.sacCer3
show(a_genome)

# Investigate some other accesors
organism(a_genome)
provider(a_genome)
seqinfo(a_genome)
```

# Discovering the Yeast genome

Let's continue to explore the yeast genome using the package 
BSgenome.Scerevisiae.UCSC.sacCer3 which is already installed for you.

As with other data in R, we can use head() and tail() to explore 
the yeastGenome object. We can also subset the genome by chromosome by using `$` as follows: 
object_name `$` chromosome_name. If you need the names of the chromosomes use the names() function.
Another nifty function is nchar(), used to count the number of characters in a sequence.

```{r}
# Load the yeast genome
library(BSgenome.Scerevisiae.UCSC.sacCer3)

# Assign data to the yeastGenome object
yeastGenome <- BSgenome.Scerevisiae.UCSC.sacCer3

# Get the head of seqnames and tail of seqlengths for yeastGenome
head(seqnames(yeastGenome))
tail(seqlengths(yeastGenome))

# Print chromosome M, alias chrM
yeastGenome$chrM

# Count characters of the chrM sequence
nchar(yeastGenome$chrM)
```

## Partitioning the Yeast genome

Genomes are often big, but interest usually lies in specific 
regions of them. Therefore, we need to subset a genome by extracting 
parts of it. To pick a sequence interval use getSeq() and specify the
name of the chromosome, and the start and end of the sequence interval:

getSeq(yeastGenome, names = "chrI", start = 100, end = 150)
Notice that names is optional; if not specified, it will return all 
chromosomes. The parameters start and end are also optional and, if 
not specified, will take the default values 1 and the length of the sequence respectively.

* Congratulations, this might have been the beginning of your first sequence analysis!
```{r}
# Load the yeast genome
#library(BSgenome.Scerevisiae.UCSC.sacCer3)

# Assign data to the yeastGenome object
#yeastGenome <- BSgenome.Scerevisiae.UCSC.sacCer3

# Get the first 30 bases of each chromosome
getSeq(yeastGenome, start = 1, end = 30)
```


# Partitioning the Yeast genome

Genomes are often big, but interest usually lies in specific 
regions of them. Therefore, we need to subset a genome by extracting 
parts of it. To pick a sequence interval use getSeq() and specify the
name of the chromosome, and the start and end of the sequence interval:

getSeq(yeastGenome, names = "chrI", start = 100, end = 150)
Notice that names is optional; if not specified, it will return all 
chromosomes. The parameters start and end are also optional and, if
not specified, will take the default values 1 and the length of the sequence respectively.

* The University of California, Santa Cruz (UCSC) Genome Browser
has made available the most genomes for BSgenome, totaling 74 of various species!


## Exploring the Zika virus sequence
It's your turn to explore the Zika virus genome, 
which has been loaded in your workspace as zikaVirus. 
The genome was downloaded from NCBI and you can apply 
Biostrings functions to learn more about it.

Start by checking the alphabet of the sequence.

alphabet() # Shows the letters included in the sequence
alphabetFrequency() # Shows the counts per letter
Remember from the video that each alphabet corresponds
to a specific biostring container, and each alphabet 
usually has extra code letters and symbols.

```{r}
library(Biostrings)
zikaVirus <-data.table::fread("zika_genomic.fa.txt")
zikaVirus <- zikaVirus$`>NC_012532.1 Zika virus isolate ZIKV/Monkey/Uganda/MR766/1947, complete genome`
zikaVirus <- DNAString(zikaVirus)
```

